---
title: "Homework 2"
author: Emily Bamforth
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r trashwheel_df}
trashwheel_df =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data!

```{r precip_2018}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

```{r precip_2017}
precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r precip_combined}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

```{r stats}
median(pull(trashwheel_df, sports_balls))

mean(pull(precip_2017, total))

mean(pull(precip_2018, total))
```

This dataset contains information about Mr. Trashwheel, a trash collector in Baltimore, MD, including the `year`, `month`, `weight_tons` of dumpsters filled, and the types of trash collected. There are `r nrow(trashwheel_df)` rows in the final Mr. Trashwheel dataset. The median number of `sports_balls` in a dumpster in 2017 was `8` balls. We also have information about precipitation by `year` and by `month`. For precipitation across 2017 and 2018, we have `r nrow(precip_df)` rows in the final dataset. The average rainfall in 2017 was `2.74` inches; in 2018, the average rainfall was `5.86` inches. 

## Problem 2

Read data.

```{r station_df}
station_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:ada) %>% 
  select(-exit_only, -staffing, -staff_hours) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

This dataset looks at the entrances and exits of NYC subway stations. There are `r nrow(station_df)` rows and `r ncol(station_df)` columns in the dataset. The dataset includes information about routes served, entrance types and whether they are ADA-compliant. So far we have read the data, cleaned the variable names using `clean_names`, selected the variables we wanted to look at and converted `entry` into a logical variable. However, this dataset is not yet tidy, since there is not one observation per row. Instead, we have multiple observations for each subway station.

```{r}
distinct_stations = station_df %>% 
  distinct(line, station_name) %>%
  nrow()

ada_compliant_stations = station_df %>% 
  distinct(line, station_name, .keep_all = TRUE) %>%
  filter(ada) %>%
  nrow()

no_vending = station_df %>% 
  distinct(line, station_name, .keep_all = TRUE) %>%
  filter(vending == "NO") %>%
  nrow()

no_vending_with_entry = station_df %>% 
  distinct(line, station_name, .keep_all = TRUE) %>%
  filter(vending == "NO" & entry == TRUE) %>%
  nrow()
```

In this dataset, there are 465 distinct stations. 84 stations are ADA-compliant. Out of 9 stations without vending, 5 allow entrance. 
